---
title: "Machine Learning"
---
# Load the necessary packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(tidymodels)
library(broom.mixed)
library(dotwhisker)
library(Metrics)
library(yardstick)
library(corrplot)
library(glmnet)
library(ranger)
```

# Set up

```{r}
##Set seed
rngseed <- 1234
set.seed(rngseed)

##Load in dsts
mavo_final <- read_rds("ml-models-exercise/mavo_final.rds")

##Recoding RACE variable
mavo_final$RACE <- recode(mavo_final$RACE, `7` = 3, `88` = 3, `1` = 1, `2` = 2)

mavo_final <- mavo_final %>% mutate(RACE = factor(RACE))

```

# Pairwise Correlation 

```{r}
##Pairwise correlation for continuous variables, also need to remove Y because its the outcome
continuous_mavo <- mavo_final %>% select(where(is.numeric)) %>% select(-Y)

cor_mat <- cor(continuous_mavo)

corrplot(cor_mat, method = "color")

```

As expected, some things (like weight and hieght) are correlated, but not too strongly, so colinearity shouldnt be a problem. 

# Feature Engineering

```{r}
##Add BMI column (Best guess looking at heights and weights, I think weights is currently in kg and heights currently in meters), formula for BMI = weight(kg) / (height(m))^2
mavo_final <- mavo_final %>% mutate(BMI = WT / (HT*HT))

```

# First Model - Linear Regression

```{r}
lm_mavo <- linear_reg() %>% set_engine("lm")
lm_mavo_fit <- lm_mavo %>% fit(Y ~ DOSE + AGE + SEX + RACE + WT + HT, data = mavo_final)
tidy(lm_mavo_fit)
tidy(lm_mavo_fit) %>% dwplot()
mean_pred_mavo <- predict(lm_mavo_fit, new_data = mavo_final)
conf_int_pred_mavo <- predict(lm_mavo_fit, new_data = mavo_final, type = "conf_int")
predict_mavo <- mavo_final %>% bind_cols(mean_pred_mavo) %>% 
    bind_cols(conf_int_pred_mavo)
metrics_mavo <- predict_mavo %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_mavo)

#Make observed vs predicted plot
predict_model1 <- predict_mavo %>% select(Y, .pred) %>%
    mutate(model = "Model 1: LM")

##Plot predicted vs observed
ggplot(predict_model1, aes(x = Y, y = .pred, color = model, shape = model)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(x = "Observed values", y = "Predicted values", title = "Observed vs Predicted Values")
```

RMSE 592. 

# Second Model - LASSO

```{r}
lasso_mavo <- linear_reg(penalty = 0.1, mixture = 1) %>% set_engine("glmnet")
lasso_mavo_fit <- lasso_mavo %>% fit(Y ~ DOSE + AGE + SEX + RACE + WT + HT, data = mavo_final)
tidy(lasso_mavo_fit)
tidy(lasso_mavo_fit) %>% dwplot()
predict_mavo2 <- predict(lasso_mavo_fit, new_data = mavo_final) %>% bind_cols(mavo_final %>% select(Y))
metrics_mavo2 <- predict_mavo2 %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_mavo2)

#Make observed vs predicted plot
predict_model2 <- predict_mavo2 %>% select(Y, .pred) %>%
    mutate(model = "Model 2: LASSO")

##Plot predicted vs observed
ggplot(predict_model2, aes(x = Y, y = .pred, color = model, shape = model)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(x = "Observed values", y = "Predicted values", title = "Observed vs Predicted Values")
```

RMSE is 592 again, as expected, it is similar to the first model. 

# Third Model - Random Forest

```{r}
rf_mavo <- rand_forest() %>% set_engine("ranger", seed = rngseed) %>% set_mode("regression")
rf_mavo_fit <- rf_mavo %>% fit(Y ~ DOSE + AGE + SEX + RACE + WT + HT, data = mavo_final)
predict_mavo3 <- predict(rf_mavo_fit, new_data = mavo_final) %>% bind_cols(mavo_final %>% select(Y))
metrics_mavo3 <- predict_mavo3 %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_mavo3)

#Make observed vs predicted plot
predict_model3 <- predict_mavo3 %>% select(Y, .pred) %>%
    mutate(model = "Model 3: Random Forest")

##Plot predicted vs observed
ggplot(predict_model3, aes(x = Y, y = .pred, color = model, shape = model)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(x = "Observed values", y = "Predicted values", title = "Observed vs Predicted Values")
```

RMSE is 379, MADA website said it was supposed to be around 362, so this is pretty close. 
Also as expected, the random forest model has values closest to the line compared to the other two models. 

# Tuning Models - Without CV (BAD PRACTICE)

```{r}
##Start with LASSO Model
recipe_all <- recipe(Y ~ ., data = select(mavo_final, where(is.numeric)))

lasso_grid <- tibble(penalty = 10^seq(-5, 2, length.out = 50))

lasso_tune <- linear_reg(penalty = tune(), mixture = 1) %>% set_engine("glmnet")

lasso_tune_wf <- workflow() %>% add_recipe(recipe_all) %>% add_model(lasso_tune)

app_resamples <- apparent(mavo_final)

lasso_tune_res <- tune_grid(lasso_tune_wf, resamples = app_resamples, grid = lasso_grid, metrics = metric_set(yardstick::rmse))

autoplot(lasso_tune_res) ###Error saying "Only one observation per metric was present. Unable to create meaningful plot."

##Now do the Random Forest Model
rf_tune <- rand_forest(trees = 300, mtry = tune(), min_n = tune()) %>% set_engine("ranger", seed = rngseed) %>% set_mode("regression")

rf_tune_wf <- workflow() %>% add_recipe(recipe_all) %>% add_model(rf_tune)

rf_grid <- grid_regular(mtry(range = c(1, 7)), min_n(range = c(1, 21)), levels = 7)

rf_tune_res <- tune_grid(rf_tune_wf, resamples = app_resamples, grid = rf_grid, metrics = metric_set(yardstick::rmse))

autoplot(rf_tune_res) ###Error saying "Only one observation per metric was present. Unable to create meaningful plot."
```

Okay the autoplot function gave me the same error twice and google and chatgpt both are telling me that using CV will fix this problem, which I think is going to be the purpose of the next part, so for now I will leave this as is?

# Tuning with CV

```{r}
set.seed(rngseed)
cv_resamples <- vfold_cv(mavo_final, v = 5, repeats = 5)

##LASSO CV tuning first
lasso_cv_res <- tune_grid(lasso_tune_wf, resamples = cv_resamples, grid = lasso_grid, metrics = metric_set(yardstick::rmse))

autoplot(lasso_cv_res)

##Now Random Forest CV tuning
rf_cv_res <- tune_grid(rf_tune_wf, resamples = cv_resamples, grid = rf_grid, metrics = metric_set(yardstick::rmse))

autoplot(rf_cv_res)

collect_metrics(lasso_cv_res)
collect_metrics(rf_cv_res)

```

As expected, the RMSE for LASSO does indeed look better than the RMSE for the Random Forest. 