---
title: "Fitting Exercise"
---
# Load the necessary packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(tidymodels)
library(broom.mixed)
library(dotwhisker)
library(Metrics)
library(yardstick)
```

# Data Processing and Exploration 

```{r}
##Read in datav
mavo <- read_csv("Mavoglurant_A2121_nmpk.csv")

##DV vs TIME plot 
ggplot(mavo, aes(x = TIME, y = DV, group = ID)) + 
    geom_line() +
    geom_point() +
    facet_wrap(~ DOSE, scales = "fixed")

##Removing OCC = 2
mavo <- mavo %>% filter(OCC == 1)

##Compute sum of DV for each individual, Exclude TIME = 0
mavo_exclude <- mavo %>% filter(TIME != 0) %>%
    group_by(ID) %>%
    summarize(Y = sum(DV, na.rm = TRUE))

##Only include TIME == 0
mavo_include <- mavo %>% filter(TIME == 0) 

##Join the two dataframes 
mavo_merged <- full_join(mavo_exclude, mavo_include, by = "ID")

##Convert RACE and SEX to factors, keeps only: Y, DOSE, AGE, SEX, RACE, WT, HT
mavo_final <- mavo_merged %>% mutate(RACE = factor(RACE), SEX = factor(SEX)) %>%
    select(Y, DOSE, AGE, SEX, RACE, WT, HT)

```

```{r}
write_rds(mavo_final, file = "mavo_final.rds")
```

# More exploring

```{r}
##Just so basic scatterplots and box plots
ggplot(mavo_final, aes(x = AGE, y = Y)) + 
    geom_point()    #looks like no relationship?

ggplot(mavo_final, aes(x = SEX, y = Y)) + 
    geom_boxplot()  #they don't look too different?

ggplot(mavo_final, aes(x = RACE, y = Y)) + 
    geom_boxplot()  #they don't look too different?

ggplot(mavo_final, aes(x = WT, y = Y)) + 
    geom_point()    #looks like no relationship?

ggplot(mavo_final, aes(x = HT, y = Y)) + 
    geom_point()    #looks like no relationship?

ggplot(mavo_final, aes(x = DOSE, y = Y, group = DOSE)) + 
    geom_boxplot()  #this doesn't surprise me, looks proportional?
```

# Model Fitting 

```{r}
##Y as outcome, DOSE as predictor
lm_yd <- linear_reg()
lm_yd_fit <- lm_yd %>% fit(Y ~ DOSE, data = mavo_final)
tidy(lm_yd_fit)
tidy(lm_yd_fit) %>% dwplot()
mean_pred_yd <- predict(lm_yd_fit, new_data = mavo_final)
conf_int_pred_yd <- predict(lm_yd_fit, new_data = mavo_final, type = "conf_int")
predict_yd <- mavo_final %>% bind_cols(mean_pred_yd) %>% 
    bind_cols(conf_int_pred_yd)
ggplot(predict_yd, aes(x = DOSE)) + 
    geom_point(aes(y = .pred)) + 
    geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper))
metrics_yd <- predict_yd %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) #this part I had to get help from chatgpt to figure out
print(metrics_yd)

```

The RMSE for this model is 666.46, meaning it's predictions of Y are typically about 666.46 units away from the true Y measured. 
The R-squared is 0.516, meaning the explains only 51.6% of the variation in Y. Typically, i would say that this means it's not a great model, but I don't even know anymore.  

```{r}
##Y as outcome, all variables as predictors
lm_y_all <- linear_reg()
lm_y_all_fit <- lm_y_all %>% fit(Y ~ DOSE + AGE + SEX + RACE + WT + HT, data = mavo_final)
tidy(lm_y_all_fit)
tidy(lm_y_all_fit) %>% dwplot()
mean_pred_y_all <- predict(lm_y_all_fit, new_data = mavo_final)
conf_int_pred_y_all <- predict(lm_y_all_fit, new_data = mavo_final, type = "conf_int")
predict_y_all <- mavo_final %>% bind_cols(mean_pred_y_all) %>% 
    bind_cols(conf_int_pred_y_all)
metrics_y_all <- predict_y_all %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_y_all)

```

The RMSE for this model is 590.85, meaning it's predictions of Y are typically about 590.85 units away from the true Y measured. 
The R-squared is 0.619, meaning the explains only 61.9% of the variation in Y. Typically, i would say that this means it's not a great model, but it does seem to explain more and have closer predictions that the previous model, which makes sense because this one account for the effects of all possible predictors. 

```{r}
##SEX as outcome, DOSE as predictor
log_sex_dose <- logistic_reg() %>% set_engine("glm")
log_sex_dose_fit <- log_sex_dose %>% fit(SEX ~ DOSE, data = mavo_final)
tidy(log_sex_dose_fit)
tidy(log_sex_dose_fit) %>% dwplot()
class_pred_sex_dose <- predict(log_sex_dose_fit, new_data = mavo_final, type = "class")
prob_pred_sex_dose <- predict(log_sex_dose_fit, new_data = mavo_final, type = "prob")
predict_sex_dose <- mavo_final %>% bind_cols(class_pred_sex_dose) %>%
    bind_cols(prob_pred_sex_dose)
event_class <- levels(mavo_final$SEX)[2]  
event_prob_col <- rlang::sym(paste0(".pred_", event_class))
metrics_sex_dose <- yardstick::metric_set(yardstick::accuracy, yardstick::roc_auc)(predict_sex_dose, truth = SEX, estimate = .pred_class, !!event_prob_col, event_level = "second") #this part I had to get help from chatgpt to figure out
print(metrics_sex_dose)

```

The model’s accuracy is 0.867, meaning 86.7% of observations were classified correctly. 
The ROC AUC is 0.592, indicating a pretty weak ability to distinguish between the two SEX categories, which is interesting. 

```{r}
##SEX as outcome, all variables as predictors
log_sex_all <- logistic_reg() %>% set_engine("glm")
log_sex_all_fit <- log_sex_dose %>% fit(SEX ~ DOSE + AGE + Y + RACE + WT + HT, data = mavo_final)
tidy(log_sex_all_fit)
tidy(log_sex_all_fit) %>% dwplot()
class_pred_sex_all <- predict(log_sex_all_fit, new_data = mavo_final, type = "class")
prob_pred_sex_all <- predict(log_sex_all_fit, new_data = mavo_final, type = "prob")
predict_sex_all <- mavo_final %>% bind_cols(class_pred_sex_all) %>%
    bind_cols(prob_pred_sex_all)
event_class <- levels(mavo_final$SEX)[2]  
event_prob_col <- rlang::sym(paste0(".pred_", event_class))
metrics_sex_all <- yardstick::metric_set(yardstick::accuracy, yardstick::roc_auc)(predict_sex_all, truth = SEX, estimate = .pred_class, !!event_prob_col, event_level = "second")
print(metrics_sex_all)

```

The model’s accuracy is 0.942, meaning 94.2% of observations were classified correctly. 
The ROC AUC is 0.980, indicating a good ability to distinguish between the two SEX categories, which is significantly better than the previous model, which normally I would say makes sense since we are accounting for all the predictors, but with the outcome being SEX, I'm not really sure what this means. 

# Module 10 

Set seed value

```{r}
rngseed <- 1234

```

# Data Prep

```{r}
##Remove RACE
mavo_final2 <- mavo_final %>% select(Y, DOSE, AGE, SEX, WT, HT)

##Set seed
set.seed(rngseed)

##Split sample
data_split <- initial_split(mavo_final2, prop = 0.75)
train_data <- training(data_split)
test_data  <- testing(data_split)

```

# Model 1

```{r}
##Same as previously, but using training data for fitting
##Y as outcome, DOSE as predictor
lm_yd1 <- linear_reg()
lm_yd_fit1 <- lm_yd1 %>% fit(Y ~ DOSE, data = train_data)
tidy(lm_yd_fit1)
tidy(lm_yd_fit1) %>% dwplot()
mean_pred_yd1 <- predict(lm_yd_fit1, new_data = train_data)
conf_int_pred_yd1 <- predict(lm_yd_fit1, new_data = train_data, type = "conf_int")
predict_yd1 <- train_data %>% bind_cols(mean_pred_yd1) %>% 
    bind_cols(conf_int_pred_yd1)
ggplot(predict_yd1, aes(x = DOSE)) + 
    geom_point(aes(y = .pred)) + 
    geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper))
metrics_yd1 <- predict_yd1 %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_yd1)

```

As it's supposed to be, RMSE is 703.

# Model 2 

```{r}
##Same as previously, but using training data for fitting
##Y as outcome, all variables as predictors
lm_y_all2 <- linear_reg()
lm_y_all_fit2 <- lm_y_all2 %>% fit(Y ~ DOSE + AGE + SEX + WT + HT, data = train_data)
tidy(lm_y_all_fit2)
tidy(lm_y_all_fit2) %>% dwplot()
mean_pred_y_all2 <- predict(lm_y_all_fit2, new_data = train_data)
conf_int_pred_y_all2 <- predict(lm_y_all_fit2, new_data = train_data, type = "conf_int")
predict_y_all2 <- train_data %>% bind_cols(mean_pred_y_all2) %>% 
    bind_cols(conf_int_pred_y_all2)
metrics_y_all2 <- predict_y_all2 %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_y_all2)

```

As it's supposed to be, RMSE is 627.

# Null Model 

```{r}
##Null model, but using training data for fitting
##Y as outcome, NO predictors
null <- null_model() %>% set_mode("regression")
null_fit <- null %>% fit(Y ~ ., data = train_data)
predict_null <- predict(null_fit, new_data = train_data) %>% 
    bind_cols(train_data %>% select(Y))
metrics_null <- predict_null %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% "rmse") 
print(metrics_null)

```

As it's supposed to be, RMSE is 948. 

# 10-fold Cross Validation

```{r}
##Set seed again
set.seed(rngseed)
##Cross-validation folds
cv_folds <- vfold_cv(train_data, v = 10)

```

# CV for Model 1 

```{r}
cv_model1 <- fit_resamples(lm_yd1, Y ~ DOSE, resamples = cv_folds)
cv_model1_metrics <- collect_metrics(cv_model1) 
print(cv_model1_metrics)

```

RMSE is now 691, different from before. 

# CV for Model 2

```{r}
cv_model2 <- fit_resamples(lm_y_all2, Y ~ DOSE + AGE + SEX + WT + HT, resamples = cv_folds)
cv_model2_metrics <- collect_metrics(cv_model2) 
print(cv_model2_metrics)

```

RMSE is 646, different from before.

# CV for Null Model 

```{r}
cv_model_null <- fit_resamples(null, Y ~ ., resamples = cv_folds)
cv_model_null_metrics <- collect_metrics(cv_model_null) 
print(cv_model_null_metrics)
```

RMSE is 933, different from before. 

# Repeat with new seed

```{r}
set.seed(999)

cv_folds_alt <- vfold_cv(train_data, v = 10)

cv_model1_alt <- fit_resamples(lm_yd1, Y ~ DOSE, resamples = cv_folds_alt)
cv_model1_alt_metrics <- collect_metrics(cv_model1_alt) 
print(cv_model1_alt_metrics)

cv_model2_alt <- fit_resamples(lm_y_all2, Y ~ DOSE + AGE + SEX + WT + HT, resamples = cv_folds_alt)
cv_model2_alt_metrics <- collect_metrics(cv_model2_alt) 
print(cv_model2_alt_metrics)

cv_model_null_alt <- fit_resamples(null, Y ~ ., resamples = cv_folds_alt)
cv_model_null_alt_metrics <- collect_metrics(cv_model_null_alt) 
print(cv_model_null_alt_metrics)

```

Now changed to 705, 648, and 933. 

# Part 2 - for this part, I got behind and so did this module late, I think I was supposed to contribute to MR repository, but because I was late, MR just did part 2 for their own work, so I will do the same here to not bother anyone else with my lateness. 

# Model Predictions 

```{r}
##Create model predictions
predict_model1 <- predict_yd1 %>% select(Y, .pred) %>%
    mutate(model = "Model 1: DOSE")

predict_model2 <- predict_y_all2 %>% select(Y, .pred) %>%
  mutate(model = "Model 2: All predictors")

predict_null_model <- predict_null %>% select(Y, .pred) %>%
  mutate(model = "Null model")

##Make Prediction Dataframe
predicted_df <- bind_rows(predict_model1, predict_model2, predict_null_model)

##Plot predicted vs observed
ggplot(predicted_df, aes(x = Y, y = .pred, color = model, shape = model)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(x = "Observed values", y = "Predicted values", title = "Observed vs Predicted Values")
```

As expected, model 2 is the best, but still lots of scatter. 

# Plot residuals 

```{r}
residuals_df <- predict_y_all2 %>% mutate(residual = .pred - Y)
max_abs_resid <- max(abs(residuals_df$residual))

ggplot(residuals_df, aes(x = .pred, y = residual)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_cartesian(ylim = c(-max_abs_resid, max_abs_resid)) +
  labs(x = "Predicted values", y = "Residuals (predicted - observed)", title = "Residual Plot for Model 2")

```

As expected, there are more and higher negative values that positive ones. 

# Model Predictions and Uncertainty

```{r}
##Reset seed
set.seed(rngseed)

##Create bootstrap
dat_bs <- bootstraps(train_data, times = 100)

##Predictions - definitely had to ask chatgpt for some help here, I was lost on it for a little while 
pred_bs <- matrix(NA, nrow = 100, ncol = nrow(train_data))
for (i in 1:100) {
  dat_sample <- rsample::analysis(dat_bs$splits[[i]])
  fit_bs <- linear_reg() %>%
    set_engine("lm") %>%
    fit(Y ~ DOSE + AGE + SEX + WT + HT, data = dat_sample)
  pred_bs[i, ] <- predict(fit_bs, train_data)$.pred}
preds <- pred_bs |> apply(2, quantile,  c(0.055, 0.5, 0.945)) |>  t() %>% as.data.frame()

##Observed and original estimates?
colnames(preds) <- c("lower", "median", "upper")
preds <- preds %>% mutate(observed = train_data$Y, point_est = predict_y_all2$.pred)

##Plot
ggplot(preds, aes(x = observed)) +
  geom_point(aes(y = point_est), color = "black", alpha = 0.6) +
  geom_point(aes(y = median), shape = 1) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(x = "Observed values", y = "Predicted values", title = "Predictions with Bootstrap Intervals") 

```

# Part 3 

# Final evaluation using test data (model 2)

```{r}
##Make predictions from fit model 2 with test data
test_predict2 <- predict(lm_y_all_fit2, new_data = test_data) %>% bind_cols(test_data) %>%
  mutate(dataset = "Test")
metrics_test2 <- test_predict2 %>% metrics(truth = Y, estimate = .pred) %>%
  filter(.metric %in% c("rmse", "rsq")) 
print(metrics_test2)

train_predict2 <- predict_model2 %>% mutate(dataset = "Train")
```

RMSE for the test data is 520, which lower than that from the training data. 

# Plot

```{r}
##Make this easier to do by combining them 
combined_predict2 <- bind_rows(train_predict2, test_predict2)
##Plot
ggplot(combined_predict2, aes(x = Y, y = .pred, color = dataset, shape = dataset)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 5000), ylim = c(0, 5000)) +
  labs(x = "Observed values", y = "Predicted values", title = "Predicted vs Observed: Training and Test Data")

```

As expected, the observed/predicted values for the test data are mixed in with the train data. 

# Discussion points 
1 - "We want to make sure that any model we have performs better than the null model. Is that the case?" - Yes, the null model RMSE is in the 900s each time, and the RMSEs for both model 1 and model 2 are lower, indicating they perform better. 

2 - "Does model 1 with only dose in the model improve results over the null model? Do the results make sense? Would you consider this model “usable” for any real purpose?" - Yes it does, I think I would call it usable in that it is bette than the null model but not quite as good as model 2 because it does not factor in the complexity of all of the other predictors as influences over the outcome. It might be good if you were specifically interested in dose as a predictor, while keeping in mind that it is not the only factor involved. 

3 - "Does model 2 with all predictors further improve results? Do the results make sense? Would you consider this model “usable” for any real purpose?" - Yes it does improve results, which makes sense because it acounts for several other predictors that can influence the results, making it a better representation of reality. I think it could be a usable model, although now that I'm thinking about it, more data would be helpful to truly evaluate how useful it and model 1 can really be. 