---
title: "Future Exercise"
---

```{r}
#load required packages
library(tidyverse)
library(dslabs)
library(dplyr)
```


```{r}
#get an overview of data structure
str(gapminder)
```


```{r}
#get a summary of data
summary(gapminder)
```


```{r}
#determine what time of object gapminder is
class(gapminder)
```


```{r}
#assign African countries to new object 
africadata <- gapminder %>% filter(continent == "Africa")

#overview of data structure
str(africadata)

#summary of data 
summary(africadata)
```


```{r}
#create new object from africadata that has only infant_mortality and life_expectancy 
africa_inf_exp <- africadata %>% select("infant_mortality", "life_expectancy")

#create new object from africadata that has only population and life_expectancy
africa_pop_exp <- africadata %>% select("population", "life_expectancy")

#structure overview of both africa_inf_exp and africa_pop_exp
str(africa_inf_exp)
str(africa_pop_exp)

#summary of both 
summary(africa_inf_exp)
summary(africa_pop_exp)
```


```{r}
#plot infant_mortality (x) vs. life_expectancy (y)
ggplot(africa_inf_exp, (aes(x = infant_mortality, y = life_expectancy))) +
  geom_point()

#plot population (x) vs. life expectancy (y)
##population in log scale so log(x) vs. (y)
###check structure
africa_pop_exp <- africa_pop_exp %>% mutate(log_population = log10(population))
str(africa_pop_exp)
ggplot(africa_pop_exp, (aes(x = log_population, y = life_expectancy))) +
    geom_point()
```

Appearance of streaks - likely represent each country over time as its population and life expextancy both slowly increase. 

```{r}
#identify where the NA values are, to avoid them in the next step 
missing_inf_years <- africadata %>% filter(is.na(infant_mortality)) %>% pull(year) %>% unique()

#new object with only year 2000 data 
africadata_2000 <- africadata %>% filter(year == 2000)

#structure of new africa_data2000
str(africadata_2000)

#plot infant_mortality (x) vs. life_expectancy (y)
ggplot(africadata_2000, (aes(x = infant_mortality, y = life_expectancy))) +
  geom_point()

##plot population (x) vs. life expectancy (y)
##population in log scale so log(x) vs. (y)
###check structure
africadata_2000 <- africadata_2000 %>% mutate(log_population = log10(population))
str(africadata_2000)
ggplot(africadata_2000, (aes(x = log_population, y = life_expectancy))) +
    geom_point()
```


```{r}
#Fit life expectancy as the outcome, and infant mortality as the predictor
fit1 <- lm(life_expectancy ~ infant_mortality, data = africadata_2000)

#Fit life expectancy as the outcome, and population as the predictor
fit2 <- lm(life_expectancy ~ population, data = africadata_2000)

#I wasn't sure if you wanted population or the log_population, so I did fit3 to be the log_population
fit3 <- lm(life_expectancy ~ log_population, data = africadata_2000)

#summary of the fits 
summary(fit1)
summary(fit2)
summary(fit3)
```

What I found here - fit1: p-value = 2.826e-08, R-squared = 0.4593 - p-value indicates significant difference (well below 0.05), but with an R-squared of 0.4593, that is ia fairly weak inverse (negative) correlation between life expectancy and infant mortality, confirmed visually with plot
fit2: p-value = 0.6159, R-squared = -0.01513 - p-value indicates no significant difference (above 0.05), no relationship between life_expectancy and population
fit3: p-vlaue = 0.4755, R-squared = -0.009755 - p-value indicates no signifiicant difference (above 0.05), no relationhsip between life expectancy and log_population