---
title: "Future Exercise"
---

```{r}
#load required packages
library(tidyverse)
library(dslabs)
library(dplyr)
```


```{r}
#get an overview of data structure
str(gapminder)
```


```{r}
#get a summary of data
summary(gapminder)
```


```{r}
#determine what time of object gapminder is
class(gapminder)
```


```{r}
#assign African countries to new object 
africadata <- gapminder %>% filter(continent == "Africa")

#overview of data structure
str(africadata)

#summary of data 
summary(africadata)
```


```{r}
#create new object from africadata that has only infant_mortality and life_expectancy 
africa_inf_exp <- africadata %>% select("infant_mortality", "life_expectancy")

#create new object from africadata that has only population and life_expectancy
africa_pop_exp <- africadata %>% select("population", "life_expectancy")

#structure overview of both africa_inf_exp and africa_pop_exp
str(africa_inf_exp)
str(africa_pop_exp)

#summary of both 
summary(africa_inf_exp)
summary(africa_pop_exp)
```


```{r}
#plot infant_mortality (x) vs. life_expectancy (y)
ggplot(africa_inf_exp, (aes(x = infant_mortality, y = life_expectancy))) +
  geom_point()

#plot population (x) vs. life expectancy (y)
##population in log scale so log(x) vs. (y)
###check structure
africa_pop_exp <- africa_pop_exp %>% mutate(log_population = log10(population))
str(africa_pop_exp)
ggplot(africa_pop_exp, (aes(x = log_population, y = life_expectancy))) +
    geom_point()
```

Appearance of streaks - likely represent each country over time as its population and life expextancy both slowly increase. 

```{r}
#identify where the NA values are, to avoid them in the next step 
missing_inf_years <- africadata %>% filter(is.na(infant_mortality)) %>% pull(year) %>% unique()

#new object with only year 2000 data 
africadata_2000 <- africadata %>% filter(year == 2000)

#structure of new africa_data2000
str(africadata_2000)

#plot infant_mortality (x) vs. life_expectancy (y)
ggplot(africadata_2000, (aes(x = infant_mortality, y = life_expectancy))) +
  geom_point()

##plot population (x) vs. life expectancy (y)
##population in log scale so log(x) vs. (y)
###check structure
africadata_2000 <- africadata_2000 %>% mutate(log_population = log10(population))
str(africadata_2000)
ggplot(africadata_2000, (aes(x = log_population, y = life_expectancy))) +
    geom_point()
```


```{r}
#Fit life expectancy as the outcome, and infant mortality as the predictor
fit1 <- lm(life_expectancy ~ infant_mortality, data = africadata_2000)

#Fit life expectancy as the outcome, and population as the predictor
fit2 <- lm(life_expectancy ~ population, data = africadata_2000)

#I wasn't sure if you wanted population or the log_population, so I did fit3 to be the log_population
fit3 <- lm(life_expectancy ~ log_population, data = africadata_2000)

#summary of the fits 
summary(fit1)
summary(fit2)
summary(fit3)
```

What I found here - fit1: p-value = 2.826e-08, R-squared = 0.4593 - p-value indicates significant difference (well below 0.05), but with an R-squared of 0.4593, that is ia fairly weak inverse (negative) correlation between life expectancy and infant mortality, confirmed visually with plot
fit2: p-value = 0.6159, R-squared = -0.01513 - p-value indicates no significant difference (above 0.05), no relationship between life_expectancy and population
fit3: p-vlaue = 0.4755, R-squared = -0.009755 - p-value indicates no signifiicant difference (above 0.05), no relationhsip between life expectancy and log_population

## Additional exploration: `dslabs::gapminder` (contributed by Alexandra Tejada-Strop)

In this section I explore the `gapminder` dataset from the `dslabs` package. I do basic exploration, some light cleaning, a few figures, and then fit a simple regression model to describe how life expectancy relates to economic output and time.

# Load packages ----
library(tidyverse)
library(dslabs)

# Load data ----
data(gapminder)  # loads a data frame called `gapminder`

# Quick look ----
glimpse(gapminder)
summary(gapminder)

# Check for missing values ----
gapminder %>%
  summarise(across(everything(), ~ sum(is.na(.))))

### Basic processing / cleaning

I create a few helpful variables:
- `gdp_per_cap`: GDP per capita
- `log_gdp_per_cap`: log10 GDP per capita (often more linear for modeling)
I also remove rows where GDP or population are missing or zero (to avoid dividing by zero and taking log of non-positive values).

gap2 <- gapminder %>%
  mutate(
    gdp_per_cap = gdp / population,
    log_gdp_per_cap = log10(gdp_per_cap)
  ) %>%
  filter(
    !is.na(gdp_per_cap),
    is.finite(log_gdp_per_cap),
    gdp_per_cap > 0,
    population > 0
  )

glimpse(gap2)

### Exploratory figures

1) Life expectancy over time by region  
2) Relationship between life expectancy and GDP per capita (log scale), colored by region  

# 1) Life expectancy over time by region
gap2 %>%
  group_by(year, region) %>%
  summarise(mean_life_exp = mean(life_expectancy), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_life_exp, color = region)) +
  geom_line() +
  labs(
    title = "Average life expectancy over time by region",
    x = "Year",
    y = "Average life expectancy"
  )

# 2) Life expectancy vs GDP per cap (log scale)
gap2 %>%
  ggplot(aes(x = log_gdp_per_cap, y = life_expectancy, color = region)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Life expectancy vs log10(GDP per capita)",
    x = "log10(GDP per capita)",
    y = "Life expectancy"
  )

### Simple statistical model

I fit a linear regression model predicting life expectancy from:
- log10(GDP per capita)
- year (to capture general time trends)
- region (to capture broad geographic differences)

This is not meant to be causal, just a simple descriptive model.

m1 <- lm(life_expectancy ~ log_gdp_per_cap + year + region, data = gap2)
summary(m1)

# A cleaner coefficient table
broom::tidy(m1) %>%
  arrange(p.value)

### Results (plain-language summary)

- The coefficient for `log_gdp_per_cap` is typically positive: higher GDP per capita is associated with higher life expectancy.
- The coefficient for `year` is typically positive: life expectancy tends to increase over time.
- Regional coefficients reflect systematic differences across regions after accounting for GDP per capita and time.

(Exact estimates may vary slightly depending on filtering and package versions.)
