The data for this exercise is the Weekly Respiratory Syncytial Virus (RSV) Vaccination Coverage among Pregnant Women by Race and Ethnicity age 18-49 in 2023-2024, gathered by National Center for Immunization and Respiratory Diseases (NCIRD), collected from the CDC's data website.Link to data: https://data.cdc.gov/Pregnancy-Vaccination/Weekly-Respiratory-Syncytial-Virus-RSV-Vaccination/g4jn-64pd/about_data
The seven columns include the week ending date, race and ethnicity, percent of pregnant women who were vaccinated, denominator being the estimated number of pregnant women, date order, race sort order, and figure ID. 

```{r}
##Load required packages
library(dplyr)
library(ggplot2)
library(tidyverse)
```

Cleaning the data

```{r}
##Read in the file

file.exists("cdcdata-exercise/Weekly_Respiratory_Syncytial_Virus_(RSV)_Vaccination_Coverage_among_Pregnant_Women_by_Race_and_Ethnicity_20260209.csv")

rsv_preg_data <- read_csv("Weekly_Respiratory_Syncytial_Virus_(RSV)_Vaccination_Coverage_among_Pregnant_Women_by_Race_and_Ethnicity_20260209.csv")

##Check on what the data looks like 
str(rsv_preg_data)

head(rsv_preg_data)

colSums(is.na(rsv_preg_data))

##Fix the Week_Ending_Date column so that ggplot knows what it is 

rsv_preg_data$Week_Ending_Date <- ymd_hms(rsv_preg_data$Week_Ending_Date)

##Also going to rename some columns just to make it easier to look at and understand

rsv_preg_data <- rsv_preg_data %>% rename(Percent_Vaxed = Percent, Total_Preg = Denominator)

##Also going to create a column using the total number pregnant and percent vaccinated to get number vaccinated 

rsv_preg_data <- rsv_preg_data %>% mutate(Number_Vax = (Percent_Vaxed/100)*Total_Preg)

##Now thinking about what percent of the total vaccinated per day does each race/ethnicity represent

rsv_preg_data <- rsv_preg_data %>% group_by(Week_Ending_Date) %>%
    mutate(Percent_of_total_vax = (Number_Vax/Number_Vax[`Race and Ethnicity` == "Overall"])*100) %>%
    ungroup()

##I'm also going to delete some columns that I don't need
rsv_preg_data <- rsv_preg_data %>% select(-Figure_ID, -`Race Sort Order`, -`Date Order`)

##Check that it all looks right 
head(rsv_preg_data)
``` 

Start exploring the data

```{r}
##Starting with looking at the percent vaccinated over time for each race/ethnicity 
ggplot(rsv_preg_data, aes(x = Week_Ending_Date, y = Percent_Vaxed, color =  `Race and Ethnicity`)) +
    geom_point() + 
    geom_line(aes(group = `Race and Ethnicity`)) +
    theme(axis.text.x = element_text(angle = 90))
```

It seems for all races and ethnicities, the percent of vaccinated mothers increases over time but then begins to plateau toward the end of the sampled dates. The overall line (the sum of all the race/ethnic groups) also follows this trend. 

```{r}
##Now looking at for each time point, what proportion of the total number vaccinated pregnant women make up each race/ethnicity. Excluding the overall, since that is the total for all groups.

ggplot(rsv_preg_data %>% filter(`Race and Ethnicity` != "Overall"), aes(x = Week_Ending_Date, y = Percent_of_total_vax, fill = `Race and Ethnicity`)) +
  geom_bar(position = "fill", stat = "identity") +
   scale_y_continuous(labels = scales::percent)
```

From this graph, the percent of vaccinated mothers from each race/ethnic group on each date is not constant. Looking at both this graph and the previous plot, it appears that the "Black, NH" Race and Ethnicity has the lowest number of pregnant mothers getting vaccinated, consistently. "Asian, NH" seems to be the fastest growing number of women getting vaccinatedand "White, NH" is most vaccinated. 

```{r}
##At this point, I want to see the overall stats, for ease of not having to search for the totals in the whole table, not the race/ethnicity specific stats, so I'm going to make a new data frame 

rsv_preg_overall <- rsv_preg_data %>% filter(`Race and Ethnicity` == "Overall") %>%
    select(-`Race and Ethnicity`, -Percent_of_total_vax)

head(rsv_preg_overall)
```

I find it interesting that, according to this table, the percent of totally pregnant women vaccinated for RSV does not exceed 17.80% in this dataset and appears to be plateuing, I wonder WHY people are not wanting the RSV vaccine. 

# This section contributed by Riley Herber

```{r}
# Creating a synthetic dataset that mimics the original real data

syn_dat <- data.frame( 
    week_ending_date = as.Date(rep(NA, 171)),
    race_and_ethnicity = numeric(171),
    percent_vaxed = numeric(171),
    total_preg = numeric(171),
    number_vax = numeric(171),
    precent_of_total_vax = numeric(171)
)

```

The following cell block was generatdeusing AI tools, which was then modified to get desired outcomes.

The original prompt was:

I want to to create a synthetic data set that represents the Weekly Respiratory Syncytial Virus (RSV) Vaccination Coverage among Pregnant Women by Race and Ethnicity age 18-49 in 2023-2024. My data structure has the structure of:
" syn_dat <- data.frame( week_ending_date = as.Date(rep(NA, 171)), race_and_ethnicity = numeric(171), percent_vaxed = numeric(171), total_preg = numeric(171), number_vax = numeric(171) ), numeric(171) " 

And I need the characteristics of: for all races and ethnicities, the percent of vaccinated mothers increases over time but then begins to plateau toward the end of the sampled dates. The overall line (the sum of all the race/ethnic groups) also follows this trend. The percent of vaccinated mothers from each race/ethnic group on each date is not constant. Looking at both this graph and the previous plot, it appears that the "Black, NH" Race and Ethnicity has the lowest number of pregnant mothers getting vaccinated, consistently. "Asian, NH" seems to be the fastest growing number of women getting vaccinatedand "White, NH" is most vaccinated. The percent of totally pregnant women vaccinated for RSV does not exceed 17.80% in this dataset and appears to be plateuing"

I wanted to put in the information given to me from the previouse analysis of the original dataset without giving AI the actual data. This was probably not the most optimally engeneered prompt, but it output the code below. 

I then modified values like the adding all of the race and ethnicity categories, modifying the number of weeks, the noise levels, and some of the logistic equations to make the graphical output more similar ot the original.
```{r}

set.seed(123)

# Weekly dates
dates <- seq.Date(as.Date("2023-01-07"), by = "week", length.out = 27)

races <- c("White, NH", "Black, NH", "Asian, NH", 
           "Hispanic", "Multiple/Other", 
           "American Indian", 
           "Native Hawaiian/Pacific Islander", 
           "Unknown")

syn_dat <- do.call(rbind, lapply(races, function(r) {
  
  t <- 1:length(dates)
  
  percent <- switch(r,
    "White, NH"  = 0.18 / (1 + exp(-0.15 * (t - 20))),
    "Black, NH"  = 0.10 / (1 + exp(-0.10 * (t - 22))),
    "Asian, NH"  = 0.2 / (1 + exp(-0.20 * (t - 18))),  # fastest growth
    "Hispanic"   = 0.150 / (1 + exp(-0.13 * (t - 21))),
    "Multiple/Other" = 0.140 / (1 + exp(-0.12 * (t - 22))),
    "American Indian" = 0.120 / (1 + exp(-0.11 * (t - 23))),
    "Native Hawaiian/Pacific Islander" = 0.130 / (1 + exp(-0.14 * (t - 21))),
    "Unknown" = 0.100 / (1 + exp(-0.09 * (t - 24)))
  )
  
  # Add week-to-week noise
  percent <- percent + rnorm(length(percent), 0, 0.002)
  percent[percent < 0] <- 0
  percent[percent > 0.178] <- 0.178
  
  total_preg <- round(runif(length(dates), 8000, 20000))
  number_vax <- round(total_preg * percent)
  percent_of_total_vax = round(100 * number_vax / total_preg, 1)
  
  data.frame(
    week_ending_date = dates,
    race_and_ethnicity = r,
    percent_vaxed = percent,
    total_preg = total_preg,
    number_vax = number_vax,
    percent_of_total_vax = percent_of_total_vax
  )
}))

```

```{r}
# Same code used as before, just changed the names to match the new synthetic data variables
ggplot(syn_dat, aes(x = week_ending_date, y = percent_vaxed, color = race_and_ethnicity)) +
  geom_point() + 
  geom_line(aes(group = race_and_ethnicity)) +
  theme(axis.text.x = element_text(angle = 90))
```


```{r}
ggplot(syn_dat, aes(x = week_ending_date, y = percent_of_total_vax, fill = race_and_ethnicity)) +
  geom_bar(position = "fill", stat = "identity") +
   scale_y_continuous(labels = scales::percent)
```

I was unable to make the Percent of total vaxenated values match the original dataset. In the original dataset, White, NH total vaxination precentage can get up to 75% in some weeks, but if I tried to edit my values to match that, the percent vaxed variable would no longer reflect the same trends seen in the original line graph figure. I think this must me because my "percent_of_total_vax" is dependent on my "percent_vax" variable. Maybe I dont understand the difference between these variables in the original dataset, because I would assume they would have to reflect one another. 

The line graph for the synthetic data matches the original data quite well, with the Asian Race having the highest percent vaxxed, the White race being second, and the Black and Native American races having some of the lowest percent vaxxed.